Code used to generate paper results : Database comparison and CT characteristics

### 2023-11-17 Paper_code-Database_CT.Rmd
Added Bruggeman analysis

```{r, warning = FALSE}
new <- CT_genes %>% 
  filter(!external_gene_name%in%common)

@@

>>>>

```{r new, warning = FALSE}
new <- CT_genes %>% 
  filter(!external_gene_name%in%common)

@@

## Other databases : minor analysis

There are ocassionnally other CT genes list that exist in litterature but there
is not always details on the pipeline, the way it is maintained or handled. 
There is usually a lack of structure.

### Bruggeman 2018

They started with the fact that most CT genes have been identified using whole 
testis, this thus don’t make them germline specific and diminishes their 
interest as targets for cancer immunotherapy. Their aim was thus to identify 
true germ-cell specific cancer genes. They thus used transcriptomes of specific 
germ cell-types present in healthy human spermatogenesis and compared it to data
from GTEx and the TCGA. This allowed them to identify 756 germ cell specific 
genes that are widely expressed in cancer. They state that "they’re certain that
their genes are true GC-genes". However, they don't validate the known 
enrichment in CT genes as their genes are distributed evenly across all 
chromosomes

The human male germ cells transcriptome they use comes from tissue samples from 
6 males, cell subtypes were isolated using a laser capture microdissection 
before RNA sequencing.

Data was available through a web application that gives now an error, and is 
also in an excel supplemental file.

```{r Bruggeman}
Bruggeman_CT <- read_excel("data/Bruggeman_suppl_data.xlsx", skip = 1,
                           sheet = "1D")

GTEX_expression(Bruggeman_CT$Gene, units = "log_TPM")
TCGA_expression(Bruggeman_CT$Gene, units = "log_TPM", tumor = "all")
testis_expression(Bruggeman_CT$Gene, cells = "all")
```

76 out of the 756 gene names aren't found in GTEx, and our data.

GTEX expression : There is clearly a lack of testis specificty. Expression in 
lymphocytes and fibroblasts is not surprising as they excluded it from their 
analysis.

TCGA expression : Most genes seems to be expressed in all tumours, which is the 
reflection of an overall expression and not an activation.

Testis cells expression : expression is indeed mainly in germ cells, even if 
some genes are expressed in somatic cells. However, lots of genes are just not
expressed in any of these cell types.

```{r Bruggeman_ours}
Bruggeman_ours <- Venn(list(Bruggeman = Bruggeman_CT$Gene,
                            CTexploreR = CT_genes$external_gene_name))
gp <- VennThemes(compute.Venn(Bruggeman_ours))
gp[["Face"]][["11"]]$fill <-  "mistyrose"
gp[["Face"]][["01"]]$fill <-  "darkseagreen1"
gp[["Face"]][["10"]]$fill <-  "lightsteelblue1"
gp[["Set"]][["Set1"]]$col <-  "paleturquoise4"
gp[["Set"]][["Set2"]]$col <-  "darkseagreen4"
gp[["SetText"]][["Set1"]]$col <-  "paleturquoise4"
gp[["SetText"]][["Set2"]]$col <-  "darkseagreen4"
plot(Bruggeman_ours, gp = gp)

Bruggeman_others <- Venn(list(Bruggeman = Bruggeman_CT$Gene,
                            Wang = Wang_CT$external_gene_name,
                            CTdatabase = CTdatabase_cleaned$external_gene_name))
plot(Bruggeman_others)
```

Despite the size of Bruggeman's list, there is a very small amount of genes
in common with CTexploreR. This is not surprising seeing the patterns of 
expression in normal and tumoral tissues above. I'll clarify the difference
below.

I however manage to recreate the Venn diagram of their fig 2, with slight 
differences probably due to the cleaning of gene names I have done in Wang and 
CTdatabase but not in Bruggeman.

```{r Bruggeman_lost}
Bruggeman_lost <- all_genes %>%
  filter(external_gene_name %in% Bruggeman_ours@IntersectionSets[["10"]])
# Missing the 76 genes

Bruggeman_lost[is.na(Bruggeman_lost$testis_specificity), ]$testis_specificity <- "not_specific"
table(Bruggeman_lost$testis_specificity)

table(Bruggeman_lost$testis_cell_type)

table(Bruggeman_lost$Higher_in_somatic_cell_type)

table(Bruggeman_lost$TCGA_category)

table(Bruggeman_lost$CCLE_category)
```

It is clear here that most genes are not testis specific. Even if they are 
indeed mostly expressed in germ cells, they are not specific to it as they are 
not testis specific. We can also see the small amount of genes activated in 
TCGA samples and CCLE cell lines.
